---
title: "Summary of bias results"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
```

## Read results 

```{r}
res <- read.table("simres.txt",header=TRUE)
# make sure simultion config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
# Alphabetically the factor level "logF1010" of the method variable
# comes before the other logF's. Move this level to the end.
res$method <- fct_relevel(res$method,"logF1010",after=Inf)
```


* We only show bias results for tests found to control
type 1 error?? SHOW ALL FOR NOW

```{r,eval=FALSE}
biased.test <- scan(file="simresBiasedtestRes.txt",what=logical())
all.configs <- with(res,paste(method,ConCaseRatio,nmatch,exptype,ncov,sep=":"))
biased.configs <- all.configs[biased.test]
for(i in 1:length(biased.configs)){
  res[all.configs==biased.configs[i],"bias"] <- NA
}
```

## CLR

* Bias of CLR always away from zero and greater than the other methods,
to the point where plots that include CLR obscure the comparison of
the other methods
    * Note: bias estimates are not trustworthy when the failure rate is high.
* Tweak the following filtering to see the effect of sample size, 
number of covariates, and effect size on bias of CLR.

```{r}
resCLR <- filter(res,method=="CLR",ConCaseRatio==1,beta==1,exptype=="binary0.1")
round(xtabs(bias ~ ncov + nmatch,data=resCLR),4)
```

## All results (other than CLR)

* Fix matching and beta for matching 1:1, 1:4, beta 0, .5, 1, 1.5
```{r}
plotres <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  #plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  droplevels(plotdat)
  p <- ggplot(plotdat,aes(x=ncov,y=bias,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge)) +
   geom_hline(yintercept=0.0, linetype="dashed", color = "grey") +
   facet_grid(nmatch~exptype,scales="free") + xlab("number of covariates")+ ylab("estimated bias") +
   #facet_grid(nmatch~.,scales="free") + xlab("number of covariates")+ ylab("estimated bias") +
   ggtitle("exposure type") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  biasdat <- plotdat[plotdat$bias-2*plotdat$bias.SE >00 | plotdat$bias+2*plotdat$bias.SE <0,c(1:6,8:9)]
  ggsave(file=filename,plot=p,device="eps")
  invisible(biasdat)
}

plotres(1,0,"bias_1to1_beta0.eps") #slight pos bias for 5 covs, nmatch=20
plotres(1,0.5,"bias_1to1_beta0.5.eps") #slight pos bias for 5 covs, nmatch=20
plotres(1,1,"bias_1to1_beta1.0.eps") #slight pos bias for 5 covs, nmatch=20
plotres(1,1.5,"bias_1to1_beta1.5.eps") #slight pos bias for 5 covs, nmatch=20

plotres(4,0,"bias_1to4_beta0.eps")
plotres(4,0.5,"bias_1to4_beta0.5.eps")
plotres(4,1,"bias_1to4_beta1.0.eps")
plotres(4,1.5,"bias_1to4_beta1.5.eps")
```


## Select results

```{r}
plotres.select <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR"& (res$nmatch==10|res$nmatch==30)&(res$exptype=="binary0.1"|res$exptype=="continuous"),]
  droplevels(plotdat)
  p <- ggplot(plotdat,aes(x=ncov,y=bias,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge)) +
   geom_hline(yintercept=0.0, linetype="dashed", color = "grey") +
   facet_grid(nmatch~exptype,scales="free") + xlab("number of covariates")+ ylab("estimated bias") +
   #facet_grid(nmatch~.,scales="free") + xlab("number of covariates")+ ylab("estimated bias") +
   ggtitle("exposure type") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  biasdat <- plotdat[plotdat$bias-2*plotdat$bias.SE >00 | plotdat$bias+2*plotdat$bias.SE <0,c(1:6,8:9)]
  ggsave(file=filename,plot=p,device="eps")
  invisible(biasdat)
}
plotres.select(1,1,"bias1.eps")
```
## Summary

* Firth is nearly unbiased always. Log-F(1,1) is bias negatively biased for continuous exposure and approximately unbiased for binary exposures. 
Log-F(2,2) is negatively biased.
* Comprehensive plots that show all the bias estimates are included in the 
Appendix.

## Targetted plots of the estimated bias

* Continuous exposure, separate results for 1:1 and 1:4 matching

* 1:1 and 1:4
    * Firth approximately unbiased with no covariates, tending
    toward a positive bias as the number of covariates increases.
    * Log-F(6,6) negatively biased at no covariates, tending towards
    unbiased as the number of covariates increases.
    * Log-F(8,8) stronger negative bias than log-F(6,6).

```{r,message=FALSE,warning=FALSE}
# filter by matching ratio and number of matched sets
pos.dodge <- 0.5
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype=="continuous",
       method=="Firth"|method=="logF66"|method=="logF88") %>%
  ggplot(aes(x=ncov,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))

filter(res,ConCaseRatio==4,beta==1,nmatch==30,exptype=="continuous",
       method=="Firth"|method=="logF66"|method=="logF88") %>%
  ggplot(aes(x=ncov,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))
```

*  Binary exposures: separate plots for 1:1 and 1:4
* 1:1 bias of Firth and log-F(1,1) small compared to 
downwardly-biased log-F(2,2).
* 1:4 bias of Firth near zero, log-F(1,1) negative bias
that increases toward unbiased as the number of covariates
increases. Log-F(2,2) has a stronger negative bias
than log_F(2,2).

```{r}
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22") %>%
  ggplot(aes(x=ncov,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") 

filter(res,ConCaseRatio==4,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22") %>%
  ggplot(aes(x=ncov,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") 
```


## Appendix

* Plots for fixed case:control ratio and exposure type. 
```{r}
filter(res,ConCaseRatio==1,exptype == "binary0.05") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.05")
filter(res,ConCaseRatio==4,exptype == "binary0.05") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.05")
##----------------
filter(res,ConCaseRatio==1,exptype == "binary0.1") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.1")
filter(res,ConCaseRatio==4,exptype == "binary0.1") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.1")
##----------------
filter(res,ConCaseRatio==1,exptype == "binary0.2") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.2")
filter(res,ConCaseRatio==4,exptype == "binary0.2") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.2")
##----------------
filter(res,ConCaseRatio==1,exptype == "continuous") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov, scale="free") +
  ggtitle("1:1 matching, continuous")
filter(res,ConCaseRatio==4,exptype == "continuous") %>%
  ggplot(aes(x=beta,y=bias,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov) +
   ylim(0,1) +
  ggtitle("1:4 matching, continuous exposure")
```

