---
title: "Summary of cover results"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
```

## Read results 

```{r}
res <- read.table("simres.txt",header=TRUE)
# make sure simulation config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
# Alphabetically the factor level "logF1010" of the method variable
# comes before the other logF's. Move this level to the end.
res$method <- fct_relevel(res$method,"logF1010",after=Inf)
```


* We only show cover results for tests found to control
type 1 error.

```{r}
biased.test <- scan(file="simresBiasedtestRes.txt",what=logical())
all.configs <- with(res,paste(method,ConCaseRatio,nmatch,exptype,ncov,sep=":"))
biased.configs <- all.configs[biased.test]
drop <- rep(FALSE,nrow(res))
for(i in 1:length(biased.configs)){
  drop[all.configs==biased.configs[i]] <- TRUE
}
res <- res[!drop,]
```

Also reduce to recommended df for the logF. Recommended is logF11 or logF22
for binary exposures and logF66 or logF88 for continuous exposures.

```{r}
keep <- 
  (res$exptype!="continuous" & (res$method %in% c("CLR","Firth","logF11","logF22"))) | 
  (res$exptype=="continuous" & (res$method %in% c("CLR","Firth","logF66","logF88")))
res <- res[keep,]
```

## Summary

* Firth is nearly unbiased always. Log-F(1,1) is negatively biased for continuous exposure and approximately unbiased for binary exposures. 
Log-F(2,2) is negatively biased.


## List of biased CI procedures that control type 1 error rate

```{r}
biascover1 <- res[res$cover + 2*res$cover.se < 0.95,c(1:6,8,12,13)]
biascover1[,7:9] <- round(biascover1[,7:9],3)
biascover2 <- res[res$cover - 2*res$cover.se > 0.95,c(1:6,8,12,13)]
biascover2[,7:9] <- round(biascover2[,7:9],3)

# View(biascover1)
# View(biascover2)
```

* Biascover1 -- under-coverage
    *  Lots of CLR: 60 simulation configurations, tends to be from bias
    of the estimator away from 0 (51/60 = 85\%) but is sometimes from
    bias towards zero.
    * **no** Firth
    * After filtering to log-F's with recommended df. still 7 logF11, 8 logF22,
    16 logF66 and 27 logF88. Log-F22 and higher df are from too much 
    shrinkage (i.e., bias toward 0).
    Log-F11 is a mix of bias toward 0 and away from 0 -- away from 0 when
    there are five covariates and log-F11 does not shrink enough (as noted
    in the bias results), but otherwise towards zero.


