---
title: "Summary of MSE results"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
```

## Read results 

```{r}
res <- read.table("simres.txt",header=TRUE)
# make sure simulation config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
# Alphabetically the factor level "logF1010" of the method variable
# comes before the other logF's. Move this level to the end.
res$method <- fct_relevel(res$method,"logF1010",after=Inf)
```


* We only show cover results for tests found to control
type 1 error.

```{r}
biased.test <- scan(file="simresBiasedtestRes.txt",what=logical())
all.configs <- with(res,paste(method,ConCaseRatio,nmatch,exptype,ncov,sep=":"))
biased.configs <- all.configs[biased.test]
drop <- rep(FALSE,nrow(res))
for(i in 1:length(biased.configs)){
  drop[all.configs==biased.configs[i]] <- TRUE
}
res <- res[!drop,]
```

Also reduce to recommended df for the logF. Recommended is logF11 or logF22
for binary exposures and logF66 or logF88 for continuous exposures.

```{r}
keep <- 
  (res$exptype!="continuous" & (res$method %in% c("CLR","Firth","logF11","logF22"))) | 
  (res$exptype=="continuous" & (res$method %in% c("CLR","Firth","logF66","logF88")))
res <- res[keep,]
```



## Summary

* MSE decreases with the amount of data (number of controls per 
matched set and/or number of matched sets) and increases slightly
with the effect size.
    * In the results shown below we fix values of these simulation
    configuration variables
* MSE decreases as we increase the prevalence of the binary exposure, and
the MSE for the continuous exposure is greater than for the binary 
exposures (refer to intepretation of exposure effect for different exposure distributions).
* Slight decrease as we decrease the number of covariates. Tie back to
events per variable.

* FIX: To compare the MSE of the different methods we choose a combination of 
exposure effect, number of controls per matched
set and number of matched sets. Do plots of estimated
MSE for select exposure types. Plot the MSE for the different
estimators as the number of covariates changes.
* Comprehensive plots that show all the MSE estimates are included in the 
Appendix.
* Due to its large bias, the MSE of CLR is so big that it obcures the
other results and so is excluded.
* Overall trends are analogous to power results: MSE lower as the amount of data (matched sets and/or matched controls) and the events-per-variable increase.
* For binary exposures, log-F22 generally has the lowest estimated MSE.
* For continuous exposures, log-F66 generally has the lowest estimated MSE
* Firth has notably higher MSE than the log-F methods when there is a 
paucity of data, but is similar to the log-F methods when there is enough 
data (e.g, 30 or more matched sets)

```{r}
plotres <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  #plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta,]
  droplevels(plotdat)
  levels(plotdat$exptype) <-c("bin 0.05","bin 0.10","bin 0.20", "cts") 
  p <- ggplot(plotdat,aes(x=ncov,y=MSE,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge)) +
   facet_grid(exptype~nmatch,scales="free") + xlab("number of covariates")+ ylab("estimated MSE") +
   ggtitle("number of matched sets") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  MSEdat <- plotdat[plotdat$MSE-2*plotdat$MSE.se >00 | plotdat$MSE+2*plotdat$MSE.se <0,c(1:6,8:9)]
  ggsave(file=filename,plot=p,device="eps")
  invisible(MSEdat)
}

plotres(1,0.5,"MSE_1to1_beta0.5.eps") 
plotres(1,1,"MSE_1to1_beta1.0.eps") 
plotres(1,1.5,"MSE_1to1_beta1.5.eps") 

plotres(4,0.5,"MSE_1to4_beta0.5.eps")
plotres(4,1,"MSE_1to4_beta1.0.eps")
plotres(4,1.5,"MSE_1to4_beta1.5.eps")
```

## Targetted plots of the estimated MSE

* Show results for 30 matched sets, $\beta=1$ and 1:1 matching with 
either a binary exposure of prevalence 0.1 or a continuous exposure.
* For the binary exposure, the MSE of Firth similar to log-$F(1,1)$
and both are higher than the MSE of log-$F(2,2)$; this finding
is generally true, except in 
very data-poor configurations (e.g., 10 matched sets)
where the MSE of Firth is higher than log-F11.
* For the continuous exposure, the MSE of Firth is higher the MSE of
log-$F(6,6)$ and log-$F(8,8)$ and the two log-F methods are similar. 
In these results log-$F(8,8)$ has slightly lower MSE than log-$F(6,6)$ but results are mixed, with log-$F(6,6)$  having lower MSE in some configurations.

```{r,message=FALSE,warning=FALSE}
# filter by matching ratio and number of matched sets
pos.dodge <- 0.5
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype=="binary0.1",
       method=="Firth"|method=="logF11"|method=="logF22") %>%
  ggplot(aes(x=ncov,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))
 ggsave(file="MSE1bin.eps",device="eps")
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype=="continuous",
       method=="Firth"|method=="logF66"|method=="logF88") %>%
  ggplot(aes(x=ncov,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))
ggsave(file="MSE1cts.eps",device="eps")
```

*  Binary exposures: separate plots for 1:1 and 1:4
* 1:1 Firth generally comparable to log-F(1,1). Log-F(2,2)
lower.
* 1:4 similar but differences less pronounced.

```{r}
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22") %>%
  ggplot(aes(x=ncov,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") 

filter(res,ConCaseRatio==4,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22") %>%
  ggplot(aes(x=ncov,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") 
```


## Appendix

* CLR has huge MSE and messes up the plots, so exclude.
* Plots for fixed case:control ratio and exposure type. 
```{r}
filter(res,ConCaseRatio==1,exptype == "binary0.05",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.05")
filter(res,ConCaseRatio==4,exptype == "binary0.05",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.05")
##----------------
filter(res,ConCaseRatio==1,exptype == "binary0.1",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.1")
filter(res,ConCaseRatio==4,exptype == "binary0.1",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.1")
##----------------
filter(res,ConCaseRatio==1,exptype == "binary0.2",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:1 matching, binary exposure w/ prevalence 0.2")
filter(res,ConCaseRatio==4,exptype == "binary0.2",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:4 matching, binary exposure w/ prevalence 0.2")
##----------------
filter(res,ConCaseRatio==1,exptype == "continuous",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:1 matching, continuous")
filter(res,ConCaseRatio==4,exptype == "continuous",method!="CLR") %>%
  ggplot(aes(x=beta,y=MSE,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se),
                 position=position_dodge(pos.dodge))  +
  #geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~ncov,scales="free") +
  ggtitle("1:4 matching, continuous exposure")
```

