---
title: "Summary of power results"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE,warning=FALSE)
```

## Load required packages

```{r}
library(tidyverse)
```

## Read results 


```{r}
res <- read.table("simres.txt",header=TRUE)
# make sure simultion config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
# Alphabetically the factor level "logF1010" of the method variable
# comes before the other logF's. Move this level to the end.
res$method <- fct_relevel(res$method,"logF1010",after=Inf)
```

* We only show power results for tests found to control
type 1 error. 
For binary exposures of any frequency (0.05, 0.1 or 0.2), Firth,
and the log-$F(m,m)$ with $m=1$ or 2 controlled type 1 error 
rates without being unduly conservative.
For continuous exposures, Firth,
and the log-$F(m,m)$ with $m=6$ or 8 controlled type 1 error 
rates without being unduly conservative.  

```{r}
biased.test <- scan(file="simresbiasedtestRes.txt",what=logical())
all.configs <- with(res,paste(method,ConCaseRatio,nmatch,exptype,ncov,sep=":"))
biased.configs <- all.configs[biased.test]
for(i in 1:length(biased.configs)){
  res[all.configs==biased.configs[i],"test.rej"] <- NA
}
keep <- 
  (res$exptype!="continuous" & (res$method %in% c("CLR","Firth","logF11","logF22"))) | 
  (res$exptype=="continuous" & (res$method %in% c("CLR","Firth","logF66","logF88")))
res <- res[keep,]
```

## Summary

* Comprehensive plots that show all the power estimates are included in the Appendix.
* Not surprisingly, power increases with the size of the exposure effect and the amount of data (number of controls per matched set and/or number of matched sets).
    * In the results shown below we fix values of these simulation
    configuration variables
* For binary exposures, power increases as we increase the exposure prevalence. Power for the continuous exposure is greater than for the binary exposures (refer to intepretation of exposure effect for different exposure distributions).
* Power decreases as we increase the number of covariates. [Tie back to
events per variable.]
* For the log-$F$ methods, power decreases as the amount of 
shrinkage increases.

* To compare the power of the different methods we choose a combination of 
exposure effect, number of controls per matched
set and number of matched sets. Do plots of estimated
power for select exposure types. Plot the power for the different
estimators as the number of covariates changes.

```{r}
plotres <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  #plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta,]
  droplevels(plotdat)
  levels(plotdat$exptype) <-c("bin 0.05","bin 0.10","bin 0.20", "cts") 
  p <- ggplot(plotdat,aes(x=ncov,y=test.rej,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
   geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
   facet_grid(exptype~nmatch,scales="free") + xlab("number of covariates")+ ylab("estimated power") +
   ggtitle("number of matched sets") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  test.rejdat <- plotdat[plotdat$test.rej-2*plotdat$test.rej.se >00 | plotdat$test.rej+2*plotdat$test.rej.se <0,c(1:6,8:9)]
  ggsave(file=filename,plot=p,device="eps")
  invisible(test.rejdat)
}

plotres(1,0.5,"power_1to1_beta0.5.eps") #slight pos test.rej for 5 covs, nmatch=20
plotres(1,1,"power_1to1_beta1.0.eps") #slight pos test.rej for 5 covs, nmatch=20
plotres(1,1.5,"power_1to1_beta1.5.eps") #slight pos test.rej for 5 covs, nmatch=20

plotres(4,0.5,"power_1to4_beta0.5.eps")
plotres(4,1,"power_1to4_beta1.0.eps")
plotres(4,1.5,"power_1to4_beta1.5.eps")
```

## Targetted plots of the estimated power

* Continuous exposure. Similar conclusions 
for different numbers of mathched sets and exposure effects.
Showing results here for 30 matched sets and exposure effect 1.
Separate plots for 1:1 and 1:4 matching.
* 1:1 matching:
The estimated power of the Firth, log-$F(6,6)$ and 
log-$F(8,8)$ tests are within simulation error of each
other except for the case of 5 nuisance covariates.
With five nuisance covariates the power of the Firth test is
lower than the two log-$F$ tests.
* 1:4 matching: Power of all three methods is within simulation
error of each other.



```{r,message=FALSE,warning=FALSE}
# filter by matching ratio and number of matched sets
pos.dodge <- 0.5
filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype=="continuous",
       method=="Firth"|method=="logF66"|method=="logF88") %>%
  ggplot(aes(x=ncov,y=test.rej,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  xlab("number of covariates")+ ylab("estimated power")
  ggsave(file="power1cts.eps",device="eps")
  
filter(res,ConCaseRatio==4,beta==1,nmatch==30,exptype=="continuous",
       method=="Firth"|method=="logF66"|method=="logF88") %>%
  ggplot(aes(x=ncov,y=test.rej,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  xlab("number of covariates")+ ylab("estimated power")
  ggsave(file="power4cts.eps",device="eps")
```

* Binary exposures: Separate plots for 1:1 and 1:4 matching.
* 1:1 Without covariates power of all methods comparable.
With one covariate power of Firth and log-F(1,1) comparable and 
power of log-F(2,2) lower. With five covariates,
power of Firth similar to log-F(1,1) for the rare
exposure and tends toward power of log-F(2,2) as 
exposure prevalence increases to 0.2. As this happens,
log-F(1,1) becomes anticonservative and is not shown.

Not sure what's going on with 5 covariates and Binary 0.2??
* 1:4 power of Firth greater than log-F(1,1) which is greater than
log-F(2,2). Differences attenuated as you increase the binary
exposure prevalence and the number of covariates.


```{r}
plotres <- filter(res,ConCaseRatio==1,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22")
levels(plotres$exptype) <- c("bin 0.05","bin 0.10","bin 0.20","cts")
  ggplot(plotres,aes(x=ncov,y=test.rej,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") +
   xlab("number of covariates")+ ylab("estimated power")

  ggsave(file="power1bin.eps",device="eps")

plotres <- filter(res,ConCaseRatio==4,beta==1,nmatch==30,exptype!="continuous",
             method=="Firth"|method=="logF11"|method=="logF22")
levels(plotres$exptype) <- c("bin 0.05","bin 0.10","bin 0.20","cts")
  ggplot(plotres,aes(x=ncov,y=test.rej,color=method)) +
     scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),position=position_dodge(pos.dodge)) +
  facet_grid(exptype~.,scales="free") +
   xlab("number of covariates")+ ylab("estimated power")

  ggsave(file="power4bin.eps",device="eps")
```

## Appendix

```{r}
plotresbin <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  #plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR" & res$exptype!="continuous",]
  droplevels(plotdat)
  p <- ggplot(plotdat,aes(x=ncov,y=test.rej,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
   geom_hline(yintercept=0.0, linetype="dashed", color = "grey") +
   facet_grid(nmatch~exptype,scales="free") + xlab("number of covariates")+ ylab("estimated power") +
   #facet_grid(nmatch~.,scales="free") + xlab("number of covariates")+ ylab("estimated power") +
   ggtitle("exposure type") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  test.rejdat <- plotdat[plotdat$test.rej-2*plotdat$test.rej.se >00 | plotdat$test.rej+2*plotdat$test.rej.se <0,c(1:6,8:9)]
  #ggsave(file=filename,plot=p,device="eps")
  invisible(test.rejdat)
}

plotresbin(1,0.5) 
plotresbin(1,1) 
plotresbin(1,1.5) 
plotresbin(4,0.5)
plotresbin(4,1)
plotresbin(4,1.5)

plotresexp <- function(ConCaseRatio,beta,filename) {
  pos.dodge <- 0.5
  #plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR",]
  plotdat <- res[res$ConCaseRatio==ConCaseRatio&res$beta==beta&res$method!="CLR" & res$exptype=="continuous",]
  droplevels(plotdat)
  p <- ggplot(plotdat,aes(x=ncov,y=test.rej,color=method)) +
   scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
   geom_point(position=position_dodge(pos.dodge),size=.3) +
   geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
   geom_hline(yintercept=0.0, linetype="dashed", color = "grey") +
   facet_grid(nmatch~exptype,scales="free") + xlab("number of covariates")+ ylab("estimated power") +
   #facet_grid(nmatch~.,scales="free") + xlab("number of covariates")+ ylab("estimated power") +
   ggtitle("exposure type") +
  theme(plot.title = element_text(hjust = 0.5, size = 11))
  print(p)
  test.rejdat <- plotdat[plotdat$test.rej-2*plotdat$test.rej.se >00 | plotdat$test.rej+2*plotdat$test.rej.se <0,c(1:6,8:9)]
  #ggsave(file=filename,plot=p,device="eps")
  invisible(test.rejdat)
}

plotresexp(1,0.5) 
plotresexp(1,1) 
plotresexp(1,1.5) 
plotresexp(4,0.5)
plotresexp(4,1)
plotresexp(4,1.5)
```
