---
title: "Summary of type 1 error rate results"
output: pdf_document
date: "2025-10-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
library(xtable)
```

## Read results 


```{r}
res <- read.table("simres.txt",header=TRUE)
# make sure simulation config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
# Alphabetically the factor level "logF1010" of the method variable
# comes before the other logF's. Move this level to the end.
res$method <- fct_relevel(res$method,"logF1010",after=Inf)
```

## Type 1 error rate control

We compared the empirical type 1 error rates of
unpenalized, Firth- \cite{HeinzePuhr10}, and log-$F$-penalized conditional logistic regression (CLR).
CLR and Firth sometimes failed so we summarize their failure rates first.
Failure rates and type 1 error plots for a subset of configurations first,
then for all configurations, to go in the Supp Methods.

### a) Failure rates of different methods

* 1:1 matching, 30 matched sets, binary exposures

```{r}
fres <- filter(res,ConCaseRatio==1,
               method=="Firth"|method=="CLR")
# only CLR and Firth ever fail
fres <- droplevels(fres)
ggplot(fres,aes(x=method,y=pfail)) + #geom_boxplot() +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + xlab("method")+ ylab("estimated failure rate") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="fail1.eps",device="eps")

#xt <- xtabs(pfail ~ nmatch + exptype,data=fres)
#xtable(cbind(xt),caption="Proportion of failed model fits for Firth-penalized CLR for different binary exposure prevalences. Results are shown for datasets comprised of 10 matched sets with 1:1 matching simulated under the null hypothesis of no exposure effect.",digits=4)
```

* * 1:1 matching, 30 matched sets, continuous exposures -- failure rates
were negligible (not shown)

* violin plots of failure rates over all simulation configurations

```{r}
fres <- filter(res,method=="CLR"|method=="Firth")
ggplot(fres,aes(x=method,y=pfail)) + geom_boxplot()
```

## b) Plot of type 1 error rates

* 1:1 matching, 30 matched sets, binary exposures


```{r}
pos.dodge <- 0.5
plotdat <- filter(res,ConCaseRatio==1,beta==0,nmatch==40,exptype!="continuous")
plotdat$exptype <- factor(droplevels(plotdat$exptype),
                          labels=c("0.05","0.10","0.20"))
ggplot(plotdat,aes(x=ncov,y=test.rej,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(~exptype)  +
  xlab("number of covariates")+ ylab("estimated type 1 error rate") +
  ggtitle("prevalence") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="t1e.eps",device="eps")
```

* Caption:
Estimated type 1 error rates for the eight methods with $\pm 2SE$
error bars for binary exposures with prevalence 0.05, 0.10
and 0.20. Results are shown for simulations of datasets with 
30 matched sets and 1:1 matching. For 


* 1:1 matching, 40 matched sets, continuous exposure


```{r}
plotdat <- filter(res,ConCaseRatio==1,nmatch==40,beta==0,exptype=="continuous")
ggplot(plotdat,aes(x=ncov,y=test.rej,color=method)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  xlab("number of covariates")+ ylab("estimated type 1 error rate")
ggsave(file="t1e_cts.eps",device="eps")
```

### Material for the discussion

Type 1 error rates and power are low when the 
binary exposure is rare. For insight into this phenomenon we start 
with the likehood ratio test from unpenalized conditional logistic regression when there are no covariates.
As shown in Appendix B, the curvature of the log-likelihood  $l(\beta_E)$
is $I_0 \frac{\exp(\beta_E)}{(1+\exp(\beta_E))^2}$ (Appendix B), where
$I_0$ is the number of matched sets that are discordant for the exposure
and $\beta_E$ is the log-OR for the exposure. When the exposure is rare,
most subjects are unexposed and $I_0$ is small; i.e., the likelihood is relatively flat. The log-likelihood-ratio and the LR statistic
is therefore small so that the test tends not to reject the null hypothesis.


## EXPERIMENTAL

* Find combinations of estimators and simulation configurations
for which the estimated type 1 error rate is 
    * within simulation error of the nominal 0.05, 
    * more than simulation error *above* the nominal 0.05 (biased tests),
    * or more than simulation error *below* the nominal 0.05 (conservative tests).


```{r}
good.test <- with(res,beta==0 & abs(test.rej-0.05) <= 2*test.rej.se)
gres <- res[good.test,
    c("method","ConCaseRatio","nmatch","exptype","ncov")]
gtab <- table(gres$method,gres$exptype)
rowlabs <- c("CLR","Firth","log-$F(1,1)$","log-$F(2,2)$","log-$F(4,4)$","log-$F(6,6)$","log-$F(8,8)$","log-$F(10,10)$")
collabs <- c("Binary, prev. 0.05","Binary, prev. 0.10","Binary, prev. 0.20","continuous, N(0,1)")
dimnames(gtab) <- list(rowlabs,collabs)
biased.test <- with(res,beta==0 & test.rej-2*test.rej.se > 0.05)
biased.test[is.na(biased.test)] <- FALSE
bres <- res[biased.test,
    c("method","ConCaseRatio","nmatch","exptype","ncov")]
btab<-table(bres$method,bres$exptype)
dimnames(gtab) <- list(rowlabs,collabs)
write(biased.test,file="simresBiasedtestRes.txt")
cons.test <- with(res,beta==0 & test.rej+2*test.rej.se < 0.05)
cons.test[is.na(cons.test)] <- FALSE
cres <- res[cons.test,
    c("method","ConCaseRatio","nmatch","exptype","ncov")]
ctab <- table(cres$method,cres$exptype)
dimnames(gtab) <- list(rowlabs,collabs)
```

* Now format the tables in LaTeX so they can be included in the 
Supplementary Materials. Cut-and-paste these into the supp mat
document and tidy up the row and column labels.

```{r}
library(xtable)
xtable(gtab,caption="Tests that control type 1 error. For each test and exposure type the table shows the count of simulation configurations (out of 30) in which the estimated type 1 error rate of the tests was within simulation error of the nominal 5%. For each exposure type the 30 simulation configurations are 2 case/control ratios by 5 numbers of matched sets by 3 numbers of covariates.")
```

```{r}
xtable(btab,caption="Biased tests. For each test and exposure type the table shows the count of simulation configurations (out of 30) in which the estimated type 1 error rate of the tests was greater than simulation error above the nominal 5%. For each exposure type the 30 simulation configurations are 2 case/control ratios by 5 numbers of matched sets by 3 numbers of covariates.")
```

```{r}
xtable(ctab,caption="Conservative tests. For each test and exposure type the table shows the count of simulation configurations (out of 30) in which the estimated type 1 error rate of the tests was more than simulation error below the nominal 5%. For each exposure type the 30 simulation configurations are 2 case/control ratios by 5 numbers of matched sets by 3 numbers of covariates.")
```

For binary0.05 all tests are conservative in all configs except
CLR which is biased for 

## Failure rates for Supp Mat

* Under the null hypothesis, only CLR and Firth ever failed. 
* Table is of exposure type by number of covariates, averaged
over the number of matched sets and matching.

```{r}
for(bb in c(0,0.5,1,1.5)){
out <- NULL
for(ee in c("binary0.05","binary0.1","binary0.2","continuous")) {
faildat1 <- filter(res,ConCaseRatio==1,beta==bb,exptype==ee) 
ftab1 <- xtabs(pfail ~ nmatch+ncov+method,data=faildat1)
out <- rbind(out,round(cbind(ftab1[,,1],NA,ftab1[,,2]),4))
}
print(xtable(out,caption=paste0("Proportion of failed model fits for unpenalized conditional logistic regression (CLR) and Firth-penalized CLR for different exposure types, numbers of covariates and numbers of matched sets. Results are for 1:1 matching and $\\beta$ = ",bb,"."),digits=4))
}
```
```{r}
for(bb in c(0,0.5,1,1.5)){
out <- NULL
for(ee in c("binary0.05","binary0.1","binary0.2","continuous")) {
faildat1 <- filter(res,ConCaseRatio==4,beta==bb,exptype==ee) 
ftab1 <- xtabs(pfail ~ nmatch+ncov+method,data=faildat1)
out <- rbind(out,round(cbind(ftab1[,,1],NA,ftab1[,,2]),4))
}
print(xtable(out,caption=paste0("Proportion of failed model fits for unpenalized conditional logistic regression (CLR) and Firth-penalized CLR for different exposure types, numbers of covariates and numbers of matched sets. Results are for 1:4 matching and $\\beta$ = ",bb,"."),digits=4))
}
```

* Classify according to failure by CLR and failure by Firth. 
What are the Firth and log-F(22) estimates? Have to do another simulation.


```{r}
indConfig <- res$ConCaseRatio==1 & res$nmatch==10 & 
  res$exptype=="binary0.05" & res$beta==0 & res$ncov==0
res[indConfig,]
```

## Plots of the estimated type 1 error rates 

* Not sure if plots of type 1 error rates are useful
but they are included here in case. 

* There are too many simulation configurations to fit
all on one plot so we split the results by the number
of matching controls.


```{r}
pos.dodge <- 0.5
filter(res,ConCaseRatio==1,beta==0) %>%
  ggplot(aes(x=ncov,y=test.rej,color=method)) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~exptype) + xlab("number of covariates")+ ylab("estimated type 1 error rate") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="t1e_1to1.eps",device="eps")
```

* Summary of 1:4 matching results:
    * blah

```{r}
filter(res,ConCaseRatio==4,beta==0) %>%
  ggplot(aes(x=ncov,y=test.rej,color=method)) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_dodge(pos.dodge),size=.3) +
  geom_linerange(aes(ymin=test.rej-2*test.rej.se,ymax=test.rej+2*test.rej.se),
                 position=position_dodge(pos.dodge)) +
  geom_hline(yintercept=0.05, linetype="dashed", color = "grey") +
  facet_grid(nmatch~exptype) + xlab("number of covariates")+ ylab("estimated type 1 error rate") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="t1e_1to4.eps",device="eps")
```
