---
title: "Convergence results"
output: pdf_document
date: "2026-07-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
library(xtable)
```

## Read results 


```{r}
res <- read.table("simres.txt",header=TRUE)
res$method[res$method=="Firth"] <- "FCLR"
res$method[res$method=="CLR"] <- "UCLR"
# make sure simulation config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}
```

## Non-convergence rates of different methods


* Over all simulation configurations (matched set size, number of matched sets, exposure type, exposure effect, and number of covariates) UCLR had the highest non-convergence rate 94.6\% of the time and FCLR had the highest non-convergence rate the remaining 5.4\% of the time.


```{r}
ConCaseRatio <- c(1,4)
nmatch <- c(10,20,30,40,50)
exptype <- c("continuous",paste0("binary",c(.05,.10,.20)))
betas <- c(0:3)/2
ncov <- c(0,1,5)
params <- expand.grid(ConCaseRatio,nmatch,exptype,betas,ncov)
names(params) <- c("ConCaseRatio","nmatch","exptype","beta","ncov")
mm <- rep(NA,nrow(params))
for(i in 1:nrow(params)) {
  ind <- (params[i,"ConCaseRatio"]==res[,"ConCaseRatio"] &
            params[i,"nmatch"]==res[,"nmatch"] &
            params[i,"exptype"]==res[,"exptype"] &
            params[i,"beta"]==res[,"beta"] &
            params[i,"ncov"] ==res[,"ncov"])
  pres <- res[ind,1:7]
  mm[i] <- which.max(pres[,"pfail"]) 
  #exp.ord <- (pres[pres$method=="UCLR","pfail"]>= pres[pres$method=="FCLR","pfail"]) &
  #              (pres[pres$method=="FCLR","pfail"]>=pres[pres$method=="logFU","pfail"])
  #if(!exp.ord) print(pres)
}
table(mm)/480
```

The maximum non-convergence rates for each method are shown below.

```{r}
tt <- res %>% group_by(method) %>% summarize(max=max(pfail),median=median(pfail))
```


* log-F methods always converged for binary exposures but not always for
the continuous exposure, though the non-convergence rates were always less than
0.1\%

* Non-convergence rate is higher the sparser the data. Data are more sparse when
    * the matched sets are smaller
    * the number of matched sets is smaller
    * the number of nuisance covariates is larger
    * for binary exposures, the exposure prevalence is lower


```{r}
fres <-  res
fres$pfail.se<- with(fres,sqrt(pfail*(1-pfail)/10000))
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=pfail,color=beta,linetype=ConCaseRatio,shape=method)) + geom_point() + geom_line() +
  geom_linerange(aes(ymin=pfail-2*pfail.se,ymax=pfail+2*pfail.se)) +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
 # scale_y_continuous(n.breaks = 4) +
  xlab("number of nuisance covariates")+ ylab("estimated non-convergence rate") +
  labs(color="exposure effect") +
  labs(linetype="num. controls")
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="SuppMatSimRes/full_nonconv.eps",device="eps")
ggsave(file="SuppMatSimRes/full_nonconv.pdf",device="pdf")
```


 Can only see the UCLR results -- redo with just UCLR
   
```{r}
fres <- filter(res,method=="UCLR")
fres$ncov <- as.numeric(as.character(fres$ncov))
fres$pfail.se<- with(fres,sqrt(pfail*(1-pfail)/10000))
ggplot(fres,aes(x=ncov,y=pfail,color=beta,linetype=ConCaseRatio)) + geom_point() + geom_line() +
  #geom_linerange(aes(ymin=pfail-2*pfail.se,ymax=pfail+2*pfail.se)) +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL),n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  #scale_y_continuous(n.breaks = 4) +
  xlab("number of nuisance covariates")+ ylab("estimated non-convergence rate") +
  labs(color="exposure effect") +
  labs(linetype="num. controls") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
   
ggsave(file="nonconv.eps",device="eps")
ggsave(file="nonconv.pdf",device="pdf")
```

