---
title: "Estimation: bias, MSE and coverage"
output: pdf_document
date: "2024-05-25"
---

```{r setup, include=FALSE, message=FALSE,warning=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load required packages

```{r}
library(tidyverse)
```

## Read results 

```{r}
res <- read.table("simres.txt",header=TRUE)
res$method[res$method=="Firth"] <- "FCLR"
res$method[res$method=="CLR"] <- "UCLR"
# make sure simultion config parameters are all factors
for(col in c("method","ConCaseRatio","nmatch","exptype","beta","ncov")) {
  res[,col] <- factor(res[,col])
}

```

```{r}
fres <-  res
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=bias,color=beta,linetype=ConCaseRatio,shape=method)) + geom_point() + geom_line() +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  xlab("number of nuisance covariates")+ ylab("estimated non-convergence rate") +
  labs(color="exposure effect") +
  labs(linetype="num. controls") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
```

Bias of CLR is well-known dwarfs the others, so eliminate and replot

```{r}
fres <- filter(res,method!="UCLR")
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=bias,color=beta,linetype=ConCaseRatio,shape=method)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL)) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  #scale_y_continuous(n.breaks = 4) +
  xlab("number of nuisance covariates")+ ylab("estimated non-convergence rate") +
  labs(color="exposure effect") +
  labs(linetype="num. controls")
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
```

Bias is closer to zero for 1:4 matching. Remove these to focus on bias under 1:1 matching
```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=bias,color=beta,shape=method)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
   geom_hline(yintercept=0, color = "grey",linewidth=2,alpha=.5) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL),n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  xlab("number of nuisance covariates")+ ylab("estimated bias") +
  labs(color="exposure effect") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
```

Bias is near 0 when the effect size is 0, so remove and replot. Also, 
results for 10 matched sets are all over the place, so exclude those too

```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1,beta!=0,nmatch!=10)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=bias,color=beta,shape=method)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
   geom_hline(yintercept=0, color = "grey",linewidth=2,alpha=.5) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL),n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  xlab("number of nuisance covariates")+ ylab("estimated bias") +
  labs(color="exposure effect") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="bias.eps",device="eps")
ggsave(file="bias.pdf",device="pdf")
```

JG bias plot

```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1,exptype!="binary0.1",beta==0.5|beta==1.5,nmatch==10|nmatch==50)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=bias,color=beta,linetype=nmatch,shape=nmatch)) + geom_point() + geom_line() +
   #geom_linerange(aes(ymin=bias-2*bias.SE,ymax=bias+2*bias.SE)) +
  facet_grid(nmatch~exptype) +
   geom_hline(yintercept=0, color = "grey",linewidth=2,alpha=.5) +
  #scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL),n.breaks=4) +
  scale_y_continuous(n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(exptype~method) + 
  xlab("number of nuisance covariates")+ ylab("estimated bias") +
  labs(color="exposure effect") + labs(shape="number of matched sets") +
  labs(linetype="number of matched sets") +
   #ggtitle("method") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="bias.eps",device="eps")
ggsave(file="bias.pdf",device="pdf")
```

## MSE

Do a plot of all MSEs, but then jump to the same subset of simulations as
in the bias plots.

```{r}
fres <-  res
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=MSE,color=beta,linetype=ConCaseRatio,shape=method)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL),n.breaks=3) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  #scale_y_continuous(n.breaks = 4) +
  xlab("number of nuisance covariates")+ ylab("estimated mean squared error") +
  labs(color="exposure effect") +
  labs(linetype="num. controls") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
```



```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1,nmatch!=10,beta==0|beta==1.5)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=MSE,color=beta,shape=method)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
  scale_y_continuous(sec.axis = sec_axis(~ . , name = "number of matched sets",breaks = NULL,labels = NULL),n.breaks=3) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(nmatch~exptype,scale="free") + 
  #scale_y_continuous(n.breaks = 4) +
  xlab("number of nuisance covariates")+ ylab("estimated mean squared error") +
  labs(color="exposure effect") +
  labs(linetype="num. controls") +
   ggtitle("exposure type") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
   ggsave(file="MSE.eps",device="eps")
ggsave(file="MSE.pdf",device="pdf")
```

JG MSE plot

```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1,exptype!="binary0.1",beta==0.5|beta==1.5,nmatch==10|nmatch==50)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=MSE,color=beta,linetype=nmatch,shape=nmatch)) + geom_point() + geom_line() +
  geom_linerange(aes(ymin=MSE-2*MSE.se,ymax=MSE+2*MSE.se)) +
  facet_grid(nmatch~exptype) +
   geom_hline(yintercept=0, color = "grey",linewidth=2,alpha=.5) +
  #scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL),n.breaks=4) +
  scale_y_continuous(n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(exptype~method) + 
  xlab("number of nuisance covariates")+ ylab("estimated MSE") +
  labs(color="exposure effect") + labs(shape="number of matched sets") +
  labs(linetype="number of matched sets") +
   #ggtitle("method") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="MSE.eps",device="eps")
ggsave(file="MSE.pdf",device="pdf")
```


Solve for estimator variance as MSE minus bias squared.

```{r}
res$var <- res$MSE - (res$bias)^2
fres <- filter(res,method!="UCLR",ConCaseRatio==1,exptype!="binary0.1",beta==0.5|beta==1.5,nmatch==10|nmatch==50)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=var,color=beta,linetype=nmatch,shape=nmatch)) + geom_point() + geom_line() +
  facet_grid(nmatch~exptype) +
   geom_hline(yintercept=0, color = "grey",linewidth=2,alpha=.5) +
  #scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL),n.breaks=4) +
  scale_y_continuous(n.breaks=4) +
  geom_point(position=position_jitter(.1),size=.3) +
  facet_grid(exptype~method) + 
  xlab("number of nuisance covariates")+ ylab("estimated variance") +
  labs(color="exposure effect") + labs(shape="number of matched sets") +
  labs(linetype="number of matched sets") +
   #ggtitle("method") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="var.eps",device="eps")
ggsave(file="var.pdf",device="pdf")
```


Coverage: now include it in estimation.

```{r}
fres <- filter(res,method!="UCLR",ConCaseRatio==1,exptype!="binary0.1",beta==0.5|beta==1.5,nmatch==10|nmatch==50)
fres$ncov <- as.numeric(as.character(fres$ncov))
ggplot(fres,aes(x=ncov,y=cover,color=beta,linetype=nmatch,shape=nmatch)) + 
  geom_point(position=position_dodge(width=.2)) + geom_line() +
   geom_linerange(aes(ymin=cover-2*cover.se,ymax=cover+2*cover.se),position=position_dodge(width=.2)) +
  facet_grid(nmatch~exptype) +
  geom_hline(yintercept=0.95, color = "grey",linewidth=2,alpha=.5) +
  #scale_y_continuous(sec.axis = sec_axis(~ . , name = "exposure type",breaks = NULL,labels = NULL),n.breaks=4) +
  scale_y_continuous(n.breaks=4) +
  facet_grid(exptype~method) + 
  xlab("number of nuisance covariates")+ ylab("estimated coverage of 95% CIs") +
  labs(color="exposure effect") + labs(shape="number of matched sets") +
  labs(linetype="number of matched sets") +
   #ggtitle("method") +
theme(plot.title = element_text(hjust = 0.5, size = 11))
ggsave(file="cover.eps",device="eps")
ggsave(file="cover.pdf",device="pdf")
```
