---
title: "Data applications: DES data"
output: pdf_document
date: "2026-07-21"
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load packages and functions

We use the following packages:

* `survival` for its `clogit()` function that performs unpenalized conditional logistic regression (UCLR)
* `coxphf` for its `coxphf()` function (Firth-penalized Cox proportional hazards)
that is called by `clogitf()` (see below) do to Firth-penalized conditional logistic regression (FCLR), and
* `tidyverse`

We also load R source files that contain code for the analyses. These 
files are loaded from local copies, but are also available on 
GitHub at https://github.com/SFUStatgen/logF_CLR.

* `clogitf.R`: contains the functions `clogitf()` for Firth-penalized conditional 
logistic regression, following a suggestion by G. Heinze, and `clogitlogF()` for log-F-penalized conditional logistic regression (logFU and logFW).
* `logFmatched.R`: contains the function `augment.logFmatched()` that 
augments an input data frame by the pseudo-matched sets needed for log-F-penalized
conditional logistic regression.


```{r}
library(survival)
library(coxphf)
library(tidyverse)
source("../SimStudy/clogitf.R") # clogitf() for Firth and clogitlogF() for log-F
source("../SimStudy/logFmatched.R") # data augmentation function 
```

## DES data

The DES data are available from the GitHub repository.

```{r}
des <- read.csv("DES.csv")
```


```{r}
des$matchedset <- des$matched.set # Our code requires that var name is matchedset
des <- select(des,matchedset,case,DES,matern.smoke)
summary(des)
xtabs(~DES + case + matern.smoke,data=des)
```

UCLR fails:

```{r}
form <- formula(case ~ DES + matern.smoke)
des.aug0 <- augment.logFmatched(form,des,m=0)
UCLR <- try(clogitlogF(form,des.aug0,maxit=500))
```

FCLR works but throws a warning unless you run it for many iterations.

```{r}
FCLR <- clogitf(form,des,maxit=500)
FCLR
confint(FCLR)
exp(confint(FCLR))
```

The uninformative prior for DES and maternal smoking is log-F(1,1). 

```{r}
des.augU <- augment.logFmatched(form,des,m=1)
logFU <- clogitlogF(form,des.augU)
logFU
confint(logFU)
exp(confint(logFU))
```

The uninformative prior for DES and maternal smoking is log-F(2,2). 

```{r}
des.augW <- augment.logFmatched(form,des,m=2)
logFW <- clogitlogF(form,des.augW)
logFW
confint(logFW)
exp(confint(logFW))
```


We also plot the profile penalized log likelihoods. For the log-F approaches
we do this by fixing the DES effect and maximizing over the maternal smoking
effect, with the DES effect included as an offset. For FCLR we use the output of
the `coxphfplot()` function.

```{r}
 profile_ll_logF <- function(beta_t,augdata) {
    offset <- beta_t * augdata$DES
    form <- formula(case ~ matern.smoke + strata(matchedset) + offset(offset))
    fit_t <- survival::clogit(form, data = augdata, weights = augdata$weights,
                              method = "efron")
    fit_t$loglik[2]
 }
profile_ll_FCLR <- function(formula,data) { # Suggestion from Heinze
  require(coxphf)
  data$start = data$matchedset
  data$stop = data$matchedset+0.1
  # Change response as done in clogit()
  newformula = formula
  newformula[[2]]=substitute(Surv(start,stop,case),list(case=formula[[2]]))
  environment(newformula) = environment(formula) 
  p <- coxphfplot(formula=newformula,data=data,profile=~DES)
  return(p)
}


```

Now run these functions to obtain the profile penalized log likelihoods
for all the methods.

```{r}
fclr <- profile_ll_FCLR(form,des)
beta_t <- fclr[,"DES"]
fclr <- fclr[,"log-likelihood"]
uclr <- logfu <- logfw <- rep(NA,length(beta_t))
for(i in seq_along(beta_t)) {
  uclr[i] <- profile_ll_logF(beta_t[i],des.aug0)
  logfu[i] <- profile_ll_logF(beta_t[i],des.augU)
  logfw[i] <- profile_ll_logF(beta_t[i],des.augW)
}
mm <- factor(rep(c("UCLR","FCLR","logFU","logFW"),each=length(beta_t)),
             levels=c("UCLR","FCLR","logFU","logFW"))
plotdat <- data.frame(logGRR=rep(beta_t,4),profll=c(uclr,fclr,logfu,logfw),
                      method=mm)
```

Finally, we can plot the curves.

```{r}
ggplot(plotdat,aes(x=logGRR,y=profll,color=method)) + geom_line() +
   scale_color_manual(values=c("black","#00AFBB","#52854C","#E7B800")) +
  xlab(expression(paste("log-OR ",beta))) + 
  ylab("Profile penalized log-likelihood") +
  geom_vline(xintercept=beta_t[which.max(fclr)],
             linetype="longdash", color="#00AFBB") +
  geom_vline(xintercept=beta_t[which.max(logfu)],
             linetype="longdash", color="#52854C") +
  geom_vline(xintercept=beta_t[which.max(logfw)],
             linetype="longdash", color="#E7B800")
ggsave(file="DESprofile.eps",device="eps")
ggsave(file="DESprofile.pdf",device="pdf")
```

## Session information for DES analysis

```{r session-info-DES}
sessionInfo()
```


