---
title: "Data applications: Type 2 diabetes trio data"
author: ""
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  pdf_document:
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE
)
library(survival)
library(coxphf)
library(tidyverse)
source("../SimStudy/clogitf.R") # clogitf() for Firth and clogitlogF() for log-F
source("../SimStudy/logFmatched.R") # data augmentation function 
```

## Purpose

This document reconstructs the case-parent trio analysis used in the manuscript.
It explains the conditional-likelihood formulation, shows how Mendelian
transmission probabilities enter as offsets in conditional logistic regression,
and fits an additive per-allele model for the Z+2 allele at the
\textit{GCK1} microsatellite locus.

The reconstructed data are based on the counts reported by
Frayling et al. (1999). The affected child is treated as the case, and the
genotypes that could have been transmitted by the same parents are treated as
matched pseudo-controls. Parental mating type defines the matched set.

## Data and notation

Let

- \(G\in\{0,1,2\}\) denote the number of copies of the Z+2 allele carried by
  the child;
- \(G_p=(g_{p1},g_{p2})\) denote the parental Z+2 allele counts;
- \(D=1\) identify the observed affected child and \(D=0\) identify a
  pseudo-control genotype.

The original draft described the genetic variants as SNPs. That is not correct
for this application: \textit{GCK1} is a microsatellite locus, and Z+2 is one
of its alleles.

The reported affected-child counts are:

```{r affected-counts}
affected_counts <- data.frame(
  mating = c("0x1", "0x1", "1x1", "1x1"),
  observed_g = c(0, 1, 0, 1),
  n = c(10, 15, 1, 1)
)

affected_counts
```

## Conditional likelihood for trio data

Conditional on parental mating type \(G_p\), the probability that the affected
child has genotype \(g\) is proportional to

\[
\exp\{z(g)^T\beta\}P(G=g\mid G_p),
\]

where \(P(G=g\mid G_p)\) is the Mendelian transmission probability. Thus the
likelihood contribution of one trio is

\[
\frac{
  \exp\{z(g)^T\beta\}P(G=g\mid G_p)
}{
  \sum_{g^\ast}
  \exp\{z(g^\ast)^T\beta\}P(G=g^\ast\mid G_p)
}.
\]

This has the form of a conditional logistic likelihood with linear predictor

\[
z(g)^T\beta+\log P(G=g\mid G_p).
\]

Therefore, the logarithm of the Mendelian transmission probability enters the
conditional logistic regression as an offset.

Because an additive constant within a matched set cancels from the conditional
likelihood, one may use either the exact log probabilities or any proportional
weights. For a \(1\times1\) mating, for example,

\[
P(G=0\mid G_p)=\frac14,\qquad
P(G=1\mid G_p)=\frac12,\qquad
P(G=2\mid G_p)=\frac14.
\]

The exact offsets are
\(\{\log(1/4),\log(1/2),\log(1/4)\}\). Subtracting the common value
\(\log(1/4)\) gives the equivalent offsets

\[
\{0,\log 2,0\}.
\]

### Additive per-allele model

The manuscript reports a single genotype relative risk for each additional
copy of the Z+2 allele. The corresponding model is

\[
z(g)=g
\]

and

\[
\operatorname{GRR}=\exp(\beta).
\]

This is different from a two-parameter model with separate genotype effects.
If one writes

\[
z(g)=
\begin{cases}
(0,0), & g=0,\\
(1,0), & g=1,\\
(1,1), & g=2,
\end{cases}
\]

then the parameters are log relative-risk increments:

\[
\exp(\beta_1)
=
\frac{P(D=1\mid G=1)}{P(D=1\mid G=0)},
\qquad
\exp(\beta_2)
=
\frac{P(D=1\mid G=2)}{P(D=1\mid G=1)}.
\]

The \(\beta_j\) themselves are log-GRRs, not GRRs. In these reconstructed data,
no affected child has genotype \(G=2\), so the two-parameter model has no information about \(\beta_2\). The single-parameter additive model is therefore the model used
below.

## Reconstructing the matched pseudo-control data

For a \(0\times1\) mating, the possible offspring genotypes are 0 and 1, each
with probability \(1/2\).

For a \(1\times1\) mating, the possible offspring genotypes are 0, 1, and 2,
with probabilities \(1/4\), \(1/2\), and \(1/4\), respectively.

Each observed affected child defines one matched set. The observed genotype is
coded as the case, and the other possible offspring genotypes are coded as
pseudo-controls.

```{r reconstruct-data}
make_set <- function(stratum, mating, observed_g) {
  if (mating == "0x1") {
    # Mendelian probabilities: P(G=0)=1/2, P(G=1)=1/2
    g <- c(0, 1)

  } else if (mating == "1x1") {
    # Mendelian probabilities: P(G=0)=1/4, P(G=1)=1/2, P(G=2)=1/4
    # Represent these probabilities by multiplicities 1:2:1.
    g <- c(0, 1, 1, 2)

  } else {
    stop("Unsupported mating type: ", mating)
  }

  case <- integer(length(g))

  # Mark exactly one copy of the observed genotype as the affected child.
  observed_rows <- which(g == observed_g)

  if (length(observed_rows) == 0L) {
    stop(
      "Observed genotype ", observed_g,
      " is incompatible with mating type ", mating, "."
    )
  }

  case[observed_rows[1L]] <- 1L

  data.frame(
    stratum = stratum,
    mating = mating,
    g = g,
    case = case
  )
}

sets <- vector("list", sum(affected_counts$n))
s <- 1L

for (i in seq_len(nrow(affected_counts))) {
  for (j in seq_len(affected_counts$n[i])) {
    sets[[s]] <- make_set(
      stratum = s,
      mating = affected_counts$mating[i],
      observed_g = affected_counts$observed_g[i]
    )
    s <- s + 1L
  }
}

trio_dat <- do.call(rbind, sets)
row.names(trio_dat) <- NULL

head(trio_dat, 12)
```

Basic checks:

```{r checks}
stopifnot(all(tapply(trio_dat$case, trio_dat$stratum, sum) == 1))
stopifnot(nlevels(factor(trio_dat$stratum)) == 27)
stopifnot(nrow(trio_dat) == 58)

with(trio_dat, table(mating, g, case))
```

## Fitting the additive conditional logistic regression

```{r fit-clr}
library(survival)

fit <- clogit(
  case ~ g + strata(stratum),
  data = trio_dat,
  method = "exact"
)

summary(fit)
```


```{r}
trio_dat$matchedset <- trio_dat$stratum
form <- formula(case ~ g)
#trio.aug0 <- augment.logFmatched(form,trio_dat,m=0) # doesn't work. Not set up for m=0
trio.aug0 <- trio_dat; trio.aug0$weights <- 1 # add weights manually
UCLR <- clogitlogF(form,trio.aug0,maxit=500)
UCLR
exp(c(UCLR$ci.lower,UCLR$ci.upper))
```

```{r}
FCLR <- clogitf(form,trio_dat,maxit=500)
FCLR
confint(FCLR)
exp(confint(FCLR))
```

The uninformative prior for the binary genotype effect is log-F(1,1). 

```{r}
trio.augU <- augment.logFmatched(form,trio_dat,m=1)
logFU <- clogitlogF(form,trio.augU)
logFU
confint(logFU)
exp(confint(logFU))
```


The weakly informative prior is log-F(2,2).

```{r}
trio.augW <- augment.logFmatched(form,trio_dat,m=2)
logFW <- clogitlogF(form,trio.augW)
logFW
confint(logFW)
exp(confint(logFW))
```




## Session information for the trio data analysis


```{r session-info-trio}
sessionInfo()
```
